---
title: "Four Parameters Categorical"
author: "Umaira"
format: html
execute: 
  echo: false
---


```{r}
#| message: false
#| warning: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(nnet)  # required for multinom()
library(gt)
library(knitr)
library(marginaleffects)
library(ggplot2)
library(plotly)
library(dplyr)
```



```{r}
 #| cache: true
nes_92 <- nes |> 
  filter(year == 1992) |> 
  select(sex, pres_vote) |> 
  drop_na() |> 
  mutate(pres_vote = as.factor(case_when(
    pres_vote == "Democrat" ~ "Clinton",
    pres_vote == "Republican" ~ "Bush",
    pres_vote == "Third Party" ~ "Perot",
  ))) 

fit_nes <- multinom_reg(engine = "nnet") |>
  fit(pres_vote ~ sex, data = nes_92)

```



```{r}
#| label: interactive-prediction-plot
#| cache: true

library(ggplot2)
library(plotly)
library(tidymodels)

plot_data <- plot_predictions(fit_nes,
                              by = "sex",
                              type = "prob",
                              draw = FALSE)

# Base ggplot
p <- plot_data |>
  ggplot(aes(x = sex, y = estimate, fill = group,
             text = paste0("Candidate: ", group,
                           "<br>Sex: ", sex,
                           "<br>Estimate: ", round(estimate, 3),
                           "<br>95% CI: [", round(conf.low, 3), ", ", round(conf.high, 3), "]"))) +
  geom_col(position = "dodge", width = 0.7) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                position = position_dodge(width = 0.7),
                width = 0.25, color = "black") +
  labs(
    title = "Predicted Voting Preferences by Sex in the 1992 U.S. Presidential Election",
    x = "Sex of Voter",
    y = "Predicted Probability",
    fill = "Candidate",
    caption = "Data Source: National Election Studies (NES), 1992"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal(base_size = 14)

# Convert to interactive plot and add subtitle as annotation
ggplotly(p, tooltip = "text") |>
  layout(
    margin = list(t = 130),  # more top space
    annotations = list(
      list(
        text = "Women were more likely to support Clinton, while men showed greater support for Perot.",
        xref = "paper", yref = "paper",
        x = 0, y = 1.09,
        showarrow = FALSE,
        font = list(size = 14)
      )
    )
  )


```

Voting behavior often varies across demographic groups, with gender frequently associated with differences in political preferences. Using data from the 1992 U.S. National Election Study, which surveyed thousands of voters across the country, we examine how voting preferences for Clinton, Bush, and Perot differed between men and women.
However, our model may be limited by potential non-representativeness in the survey data, as certain groups may be underrepresented among those who responded to the study. We model voter preference as a multinomial function of sex to capture how gender relates to support for each of the major presidential candidates. In particular, the predicted probability of voting for Clinton was approximately 52.5% for women and 42.5% for men, with a difference of 10 percentage points and a 95% confidence interval that does not include zero.


$$P(Y = k) = \frac{e^{\beta_{k0} + \beta_{k1} X_1 + \beta_{k2} X_2 + \cdots + \beta_{kn} X_n}}{\sum_{j=1}^{K} e^{\beta_{j0} + \beta_{j1} X_1 + \beta_{j2} X_2 + \cdots + \beta_{jn} X_n}}$$

with $Y \sim \text{Multinomial}(\boldsymbol{\rho})$ where $\boldsymbol{\rho} = (\rho_1, \rho_2, \ldots, \rho_K)$ are the probabilities above.


$$
\begin{aligned}
\hat{\rho}_{clinton} &= \frac{e^{0.45 - 0.25 \cdot male}}{1 + e^{0.45 - 0.25 \cdot male}} \\
\hat{\rho}_{perot}   &= \frac{e^{-0.85 + 0.42 \cdot male}}{1 + e^{-0.85 + 0.42 \cdot male}} \\
\hat{\rho}_{bush}    &= 1 - \hat{\rho}_{clinton} - \hat{\rho}_{perot}
\end{aligned}
$$


```{r}
#| message: false
#| warning: false
library(broom)
library(gt)

tidy(fit_nes, conf.int = TRUE) |>
  select(y.level, term, estimate, conf.low, conf.high) |>
  gt() |>
  tab_header(
    title = "Model Estimates: Predicting 1992 Presidential Vote by Sex"
  ) |>
  tab_source_note(
    source_note = "Source: National Election Studies (NES) 1992"
  ) |>
  fmt_number(
    columns = c(estimate, conf.low, conf.high),
    decimals = 3
  ) |>
  cols_label(
    y.level = "Outcome",
    term = "Term",
    estimate = "Estimate",
    conf.low = "Lower 95%",
    conf.high = "Upper 95%"
  )

```
